### 2023-01-18 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@29c663e2b848f627726a16f11a5eada481426451 🚀

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature.The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001), as do estimates
of population abundance (Griffiths 1998). This relationship between
sampling effort and spatial coverage and species detectability has
motivated more quantitatively robust approaches to account for error in
sampling data in many contexts: to determine if a given species is
extinct (Boakes \emph{et al.} 2015), to determine sampling design (Moore
\& McCarthy 2016), and to measure species richness across large scales
(Carlson \emph{et al.} 2020). In the context of interactions, an initial
concern was the compounding effects of limited sampling effort combined
with the amalgamation of data (across both study sites, time of year,
and taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (Martinez \emph{et al.} 1999;
McLeod \emph{et al.} 2021). Martinez \emph{et al.} (1999) showed that in
a plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another in proportion to each species'
relative-abundance. This assumes that there are no associations in
species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are common in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, ever exhibit any meaningful effect on the fitness of the
other. So, although co-occurrence is not directly indicative of an
interaction, it \emph{is} a precondition for an interaction.

>>>>

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature. The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001), as do estimates
of population abundance (Griffiths 1998). This relationship between
sampling effort, spatial coverage, and species detectability has
motivated more quantitatively robust approaches to account for error in
sampling data in many contexts: to determine if a given species is
extinct (Boakes \emph{et al.} 2015), to determine sampling design (Moore
\& McCarthy 2016), and to measure species richness across large scales
(Carlson \emph{et al.} 2020). In the context of interactions, an initial
concern was the compounding effects of limited sampling effort combined
with the amalgamation of data (across both study sites, time of year,
and taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (Martinez \emph{et al.} 1999;
McLeod \emph{et al.} 2021). Martinez \emph{et al.} (1999) showed that in
a plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another in proportion to each species'
relative-abundance. This assumes that there are no associations in
species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of these
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are common in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, ever exhibit any meaningful effect on the fitness of the
other. So, although co-occurrence is not directly indicative of an
interaction, it \emph{is} a precondition for an interaction.

### 2023-01-18 preprint.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@29c663e2b848f627726a16f11a5eada481426451 🚀

\marginpar{\vskip 1em\flushright
{\small{\bfseries Keywords}:\par
community ecology\\network ecology\\sampling effort\\species
interactions\\spatial sampling\\simulation\\}
}

>>>>

\marginpar{\vskip 1em\flushright
{\small{\bfseries Keywords}:\par
species interactions\\network ecology\\sampling effort\\spatial
ecology\\null model\\}
}
